\Ace: Growing a Statically-Typed Language Inside a Python

### 2013-11-14 ace-pldi14/ace-pldi14.tex
Fixed figures 1-5

Generic functions are represented at compile-time as instances of \verb|ace.GenericFn| and consist of an abstract syntax tree, an {active base} and an {active target}. The purpose of the \emph{decorator} on line 3 is to replace the Python function on lines 4-8 with an Ace generic function having the same syntax tree and the provided active base and active target. 
Decorators in Python are simply syntactic sugar for applying the decorator function directly to the function  being decorated \cite{python}. In other words, line 3 could be replaced by inserting the following  statement on line 9:

@@

\subsection{Metaprogramming in Ace}
Generic functions can be generated directly from ASTs as well, providing Ace with support for  straightforward metaprogramming. Listing \ref{metaprogramming} shows how to generate two more generic functions, \verb|thresh_scale| and \verb|thresh_dbl|. The latter is derived from the former using a library for manipulating Python syntax trees, \verb|astx|. In particular, its \verb|specialize| function replaces uses of the second argument of \verb|thresh_scale| with the literal \verb|2.0| (and changes the function's name), leaving a function of one argument.

@@

\begin{codelisting}
\begin{lstlisting}[style=Bash]
$ `acec listing3.py`
Hello, compile-time world!
[acec] listing3.cl successfully generated.
\end{lstlisting}
\caption{Compiling \texttt{listing\ref{mapadd5dbl}.py} using the \texttt{acec} compiler.}
\label{mapc}
\end{codelisting}
\begin{codelisting}
\lstinputlisting[style=OpenCL]{listing5.cl}
\caption{[\texttt{listing\ref{mapadd5dbl}.cl}] The OpenCL file generated by Listing \ref{mapc}.}
\label{mapout}
\end{codelisting}
As defined on Lines 5-7, \verb|main| is a \emph{generic function}. This terminology in Ace is used to indicate a function for which types have not yet been assigned to arguments. Invoking the \verb|compile| method on a generic function with a sequence of argument types invokes the active typechecking and translation mechanism we will describe in Section 3 to produce a \emph{concrete function} -- one with a single type for each argument, internal variable (such as \verb|hello|) and the return value. In Listing \ref{hello}, we name this concrete function \verb|main|, overwriting the generic function of the same name that it is derived from. The compiler does nothing apparently interesting: there are no arguments, the single internal variable takes the OpenCL string type \verb|char*| from its value and the return type is \verb|void|.

>>>>

Generic functions are represented at compile-time as instances of \verb|ace.GenericFn| and consist of an abstract syntax tree, an {active base} and an {active target}. The purpose of the \emph{decorator} on line 3 is to replace the Python function on lines 4-8 with an Ace generic function having the same syntax tree and the provided active base and active target. 
Decorators in Python are simply syntactic sugar for applying the decorator function directly to the function  being decorated \cite{python}. In other words, line 3 could be replaced by inserting the following  statement on line 9\todo{figure out how to center this}:

@@

\subsection{Metaprogramming in Ace}
Generic functions can be generated directly from ASTs as well, providing Ace with support for  straightforward metaprogramming. Listing \ref{metaprogramming} shows how to generate two more generic functions, \verb|scale| and \verb|double|. The latter is derived from the former using a library for manipulating Python syntax trees, \verb|astx|. In particular, its \verb|specialize| function replaces uses of the second argument of \verb|scale| with the literal \verb|2| (and changes the function's name), leaving a function of one argument.

@@

\begin{codelisting}
\begin{lstlisting}[style=Bash]
$ `acec listing3.py`
Hello, compile-time world!
[ace] TypeError in listing1.py (line 6, col 28): 
      'GenericFnType(negate)' does not support [].
[acec] listing3.cl successfully generated.
\end{lstlisting}
\caption{Compiling \texttt{listing\ref{mapadd5dbl}.py} using the \texttt{acec} compiler.}
\label{mapc}
\end{codelisting}
\begin{codelisting}
\lstinputlisting[style=OpenCL]{listing5.cl}
\caption{[\texttt{listing\ref{mapadd5dbl}.cl}] The OpenCL file generated by Listing \ref{mapc}.}
\label{mapout}
\end{codelisting}
As defined on Lines 5-7, \verb|main| is a \emph{generic function}. This terminology in Ace is used to indicate a function for which types have not yet been assigned to arguments. Invoking the \verb|compile| method on a generic function with a sequence of argument types invokes the active typechecking and translation mechanism we will describe in Section 3 to produce a \emph{concrete function} -- one with a single type for each argument, internal variable (such as \verb|hello|) and the return value. In Listing \ref{hello}, we name this concrete function \verb|main|, overwriting the generic function of the same name that it is derived from. The compiler does nothing apparently interesting: there are no arguments, the single internal variable takes the OpenCL string type \verb|char*| from its value and the return type is \verb|void|.
